---
title: "4 Questions"
output: html_document
date: "2024-09-12"
---
<h2>**Question #1**</h2>
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse")
library(tidyverse)
# cvs import 
major_list <- read.csv("C:/Users/tiffh/Downloads/4 Questions HW/majors-list.csv")
head(major_list)
# use str to filter DATA + STATA 
data_stat_major <- major_list %>%
  filter(str_detect(Major, "DATA") | str_detect(Major, "STATISTICS"))
# View which majors have data or stats in name 
data_stat_major
```
Management information systems and statistics,Computer programming and data 
processing, and Statistics and decision science where the three majors that 
contained data or statistics when filtered. They belong to the major category of 
business and computers & mathematics. 
<h2>**Question #2**</h2>
```{r setup2, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
# Fruit name vector 
fruit_name <- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

print(fruit_name)
# Fruit list assigned the fruit a number 
fruit_number <- c(1,2,3,4,5,6,7,8,9,10,11,12,13,14)

fruit_name <- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

fruit_list = list(fruit_number,fruit_name)
print(fruit_list)

```
I initially made a vector with just the fruit name, but you can also make a list 
just by assigning the fruit to a number.

<h2>**Question #3**</h2>
For the questions lets assume that there a vector C("q","s","t","r") which is
a string of " q s t r"

1. (.)\1\1
(.) = q
/1 = refers to the first variable 
So (.)\1\1 = qqq

2. (.)(.)\\2\\1
(.) = q
(.) = s
\\2 = refers to the second variables 
\\1 = refers to first variable 
Which means (.)(.)\\2\\1 = qssq

3. (..)\1
(..) = qs
\1 = refers to first variables 
Meaning (..)\1 = qsqs

4. (.).\\1.\\1
(.) = q
. = s
\\1 = refers to first variable 
So (.).\\1.\\1 = qsqq

5. (.)(.)(.).*\\3\\2\\1
(.) = q
(.) = s
(.) = t
.* = r
\\3 = refers to third variable 
\\2 = refers to second variable 
\\1 = refers to first variable 
So (.)(.)(.).*\\3\\2\\1 = qstrtsq

<h2>**Question #4**</h2>
```{r setup4, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(dplyr)
# create a vector of 50 states 
 
states <- c("alabama", "alaska", "arizona", "arkansas", "california", "colorado", 
            "connecticut", "delaware", "florida", "georgia", "hawaii", "idaho", 
            "illinois", "indiana", "iowa", "kansas", "kentucky", "louisiana", 
            "maine", "maryland", "massachusetts", "michigan", "minnesota", 
            "mississippi", "missouri", "montana", "nebraska", "nevada", 
            "new hampshire", "new jersey", "new mexico", "new york", 
            "north carolina", "north dakota", "ohio", "oklahoma", 
            "oregon", "pennsylvania", "rhode island", "south carolina", 
            "south dakota", "tennessee", "texas", "utah", "vermont", 
            "virginia", "washington", "west virginia", "wisconsin", "wyoming")

#Start and end with the same character
str_subset(states, "^(.)((.*\\1$)|\\1?$)")


# Contain a repeated pair of letters
str_subset("mississippi", "([A-Za-z][A-Za-z]).*\\1")
str_subset(states, "([A-Za-z][A-Za-z]).*\\1")


#Contain one letter repeated in at least three places
str_subset("alabama", "([a-z]).*\\1.*\\1")
str_subset(states, "([a-z]).*\\1.*\\1")
```

For the last question I created a vector of the 50 states as word pool. alabama,
alaska, arizona, and ohio are the four states that all begin and end with the 
same letter. And mississippi is the only state with repeating letter "ss". Then
there are 13 states(AL,AK,AR,CO,CT,IL,MA,MS,NJ,PA,
TN,VA,WV) that have one letter that repeats at least three places. 




**Reference**
Q1. https://brshallo.github.io/r4ds_solutions/14-strings.html

Q2. https://www.geeksforgeeks.org/r-lists/

Q3/4.https://jrnold.github.io/r4ds-exercise-solutions/strings.html
