---
title: "Assignment#10"
author: "Tiffany Hugh"
date: "2024-11-02"
output: html_document
---
**Introduction**

For Assignment #10, I will conduct a sentiment analysis of Grimm's Fairy Tales
by Jacob and Wilhelm Grimm, sourced from Project Gutenberg. This analysis will
apply the Bing, AFINN, NRC, and Vader sentiment lexicons to explore the
emotional tone of the text.

**Source**

Project Gutenberg offers free eBooks where I was able to download the Grimm's 
Fairy Tale as a txt file. In order to work with the text I first need to
convert the text to lowercase to standardize it.Remove punctuation, numbers, and
special characters. Which will help me tokenize the text into individual words 
for analysis. 

**Import and Tidy**
Before that all happens, I need the appropriate packages to access sentiment
lexicons, clean and manipulate text, and join sentiment scores to my text data.
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages(c("tidytext", "textdata", "vader", "dplyr", "tidyr"))
library(tidytext); library(textdata); library(vader); library(dplyr); library(tidyr)

# import txt file 
grimm_text<- readLines("C:/Users/tiffh/Downloads/Assignment#10/grimm.txt")

# convert to data frame
grimm_df <- data.frame(line = 1:length(grimm_text), text = grimm_text) %>%
  unnest_tokens(word, text)
```

**Bing Lexicon **
The Bing lexicon classifies words into positive and negative categories, 
offering a clear measure of sentiment polarity. To begin the analysis, first 
sentiment scores are extract by applying the Bing lexicon. This involves using
the inner_join function to merge the tokenized words with their corresponding
sentiment classifications, allowing quantification of positive and 
negative words in the text.
```{r setup2, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

bing_sentiments <- get_sentiments("bing")
# count sentiments in the Grimm text
grimm_sentiment <- grimm_df %>%
  inner_join(bing_sentiments) %>%
  count(index = line %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

#install.packages("ggplot2")
library(ggplot2)
# plot
ggplot(grimm_sentiment, aes(index, sentiment)) +
  geom_col(aes(fill = sentiment), show.legend = FALSE) +
  labs(title = "Sentiment Analysis of Grimm Fairy Tales",
       x = "Index",
       y = "Sentiment Score") +
  theme_minimal()
```

The plot illustrates how sentiment shifts between positive and negative 
throughout the narrative arc of the Grimm Fairy Tales. Given that the collection
comprises multiple distinct stories, the sentiment at the beginning tends to be
more negative, while the later tales exhibit a more positive sentiment. This 
variability reflects the diverse themes and emotional tones present within the
different stories in the collection.

**NRC Lexicon**

The NRC lexicon categorizes words into eight emotional categories along with 
positive or negative sentiment, offering a broad emotional profile of the text.
From the Bing sentiment and the knowlege that Grimm's tales will be more negative
sentiments. I focused on sadness, by filtering for sad words and then join the 
text with the nrc sadness words. A count is taken of the sadness words and joined
for analysis. 
```{r setup3, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

nrc_sadness <- get_sentiments("nrc") %>%
  filter(sentiment == "sadness")
# count occurrences of sadness words in text
sadness_word_counts <- grimm_df %>%
  inner_join(nrc_sadness, by = "word") %>%  
  count(word, sort = TRUE)                  

print(sadness_word_counts)

# join data with sadness words for detailed inspection
sadness_words_joined <- grimm_df %>%
  inner_join(nrc_sadness, by = "word")    

print(sadness_words_joined)

# subset for top sadness words 
top_sadness_words <- sadness_words_joined %>%
  count(word) %>%                 
  top_n(10, wt = n)               

# plot
ggplot(top_sadness_words, aes(x = reorder(word, n), y = n)) +
  geom_bar(stat = "identity", fill = "purple") + 
  coord_flip() +                                 
  labs(title = "Top 10 Sadness Words in Grimm's Fairy Tales",
       x = "Sadness Words",
       y = "Count") +                           
  theme_minimal() 
```

I joined the text data with the NRC sadness lexicon to gain deeper insights
into the words most commonly associated with the sentiment of sadness. Upon 
conducting this analysis, I discovered that the word "mother" emerged as the 
top sadness word in the dataset. To ensure accuracy, I performed a further
inspection by filtering for sadness words only, yet surprisingly, "mother" 
still ranked high.I assumed there would be more words associated with sadness
but maybe another emotion would better fit the text. 

**AFINN Lexicon**

The AFINN lexicon assigns words with numerical scores (-5 to 5) based on their
emotional intensity, helping to capture varying levels of sentiment. The first 
step is to get sentiment scores using the AFINN lexicon, and then using inner_
join function will match the words in your tokenized text with the words from
the text. 
```{r setup1, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

afinn_sentiments <- get_sentiments("afinn")

afinn_results <- grimm_df %>%
  inner_join(afinn_sentiments, by = "word") %>%
  count(word, score = value, sort = TRUE)

cat("AFINN Sentiment Analysis Results:\n")
print(head(afinn_results))


# selection of the top 10 positive and negative sentiments
top_positive_words <- afinn_results %>%
  filter(score > 0) %>%
  top_n(10, score)

top_negative_words <- afinn_results %>%
  filter(score < 0) %>%
  arrange(score) %>%  # Arrange by score in ascending order
  slice_head(n = 10) 

# plot 
positive <- ggplot(top_positive_words, aes(x = reorder(word, score), y = score, fill = "Positive")) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top Positive AFINN Sentiment Words in Grimm's Fairy Tales",
       x = "Words",
       y = "Sentiment Score") +
  scale_fill_manual(values = "gold") +
  theme_minimal()

negative <- ggplot(top_negative_words, aes(x = reorder(word, score), y = score, fill = "Negative")) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top Negative AFINN Sentiment Words in Grimm's Fairy Tales",
       x = "Words",
       y = "Sentiment Score") +
  scale_fill_manual(values = "blue") +
  theme_minimal()

print(positive)
print(negative)
```

There were many words, so I examined the top 10 for both positive and negative
sentiments. These are presented in separate plots. The top positive words are 
associated with themes of celebration, joy, and humor. In contrast, the negative
sentiments reflect more distressing themes, including violence and derogatory 
language. 

**Comparison**

I’m looking to see which lexicon from Chapter 2 works best for Grimm's Fairy 
Tales. I’ll put together a summary for each lexicon to show how well they
reflect the story's sentiments.
```{r setup5, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
# Prepare a summary for each lexicon
bing_summary <- grimm_df %>%
  inner_join(bing_sentiments, by = "word") %>%
  count(sentiment, sort = TRUE) %>%
  mutate(lexicon = "Bing")

nrc_sentiments <- get_sentiments("nrc")
nrc_summary <- grimm_df %>%
  inner_join(nrc_sentiments, by = "word") %>%
  count(sentiment, sort = TRUE) %>%
  mutate(lexicon = "NRC")

afinn_summary <- grimm_df %>%
  inner_join(afinn_sentiments, by = "word") %>%
  count(score = value, sort = TRUE) %>%
  mutate(sentiment = ifelse(score > 0, "Positive", "Negative"),
         lexicon = "AFINN") %>%
  group_by(sentiment, lexicon) %>%
  summarize(n = sum(n), .groups = 'drop')

# Combine all summaries
combined_summary <- bind_rows(bing_summary, nrc_summary, afinn_summary)

# Print the combined summary
print(combined_summary)

# Plot the comparison of sentiment counts across lexicons
ggplot(combined_summary, aes(x = lexicon, y = n, fill = sentiment)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Sentiment Lexicons on Grimm's Fairy Tales",
       x = "Lexicon",
       y = "Count") +
  theme_minimal()

```

This chart compares three different sentiment lexicons—AFINN, Bing, and NRC—used
to analyze sentiments in Grimm's Fairy Tales. Each lexicon has its strengths: 
AFINN gives an overall sentiment count, while Bing captures positive and negative 
sentiments. However, neither really dives into the specific emotions that fairy
tales often express, which can be quite complex. That’s why I believe NRC is the
best fit. It’s the most comprehensive, capturing specific emotions like anger, 
anticipation, joy, sadness, and more. Since fairy tales are full of varied 
emotions, NRC offers deeper insights, allowing for a better understanding of 
the emotional landscape in Grimm's stories. Based on this chart, NRC stands out 
as the most suitable lexicon for analyzing sentiments in Grimm's Fairy Tales 
because it covers a wide range of specific emotions, providing a richer 
emotional analysis.


**VADAR Lexicon**

As an additional layer, I will include the VADER lexicon. While it is typically
used for social media analysis, I thought it would fit Grimm's Fairy Tales well 
because its ability to capture nuanced sentiments in text can reveal the
underlying emotional tones present in these classic stories. Given the complex
characters and moral dilemmas within the tales, VADER’s focus on both positive
and negative sentiments will enhance the overall understanding of the narratives
emotional landscape.

While trying to run VADER, I encountered several difficulties that slowed down 
the process significantly. Initially, the full dataset caused long processing
times, likely due to its size and complexity, which made it challenging to
manage and analyze efficiently. Additionally, I faced warnings regarding the 
data structure, such as the message indicating that the number of items to 
replace was not a multiple of the replacement length. This suggested there 
were mismatches in the expected data format, leading to further complications.


Given these challenges, I decided to simplify my approach and run the VADER
sentiment analysis on only a sample of the data. By focusing on a smaller 
subset, I could more effectively troubleshoot any issues and achieve quicker 
results, ultimately allowing for a clearer understanding of the sentiment
dynamics present in Grimm's Fairy Tales without the overwhelming burden of
processing the entire dataset at once.

```{r setup6, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
# re-imported data so i can do it line by line instead of by word
grimm_text <- readLines("C:/Users/tiffh/Downloads/Assignment#10/grimm.txt")
grimm_df <- data.frame(text = grimm_text, stringsAsFactors = FALSE)
print(head(grimm_df))

# subset of the first 100 lines for faster processing
sample_grimm_df <- grimm_df[1:100, ]

# sentiment analysis on the sample
vader_results_sample <- vader_df(sample_grimm_df)

print(head(vader_results_sample))
colnames(vader_results_sample)

#stats on smaple
summary_statistics <- vader_results_sample %>%
  summarise(
    mean_positive = mean(pos, na.rm = TRUE),
    mean_negative = mean(neg, na.rm = TRUE),
    mean_neutral = mean(neu, na.rm = TRUE),
    mean_compound = mean(compound, na.rm = TRUE)
  )

print(summary_statistics)

#plot 
sentiment_counts <- vader_results_sample %>%
  summarise(
    Positive = sum(pos > 0, na.rm = TRUE),
    Negative = sum(neg > 0, na.rm = TRUE),
    Neutral = sum(neu > 0, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = everything(), names_to = "Sentiment", values_to = "Count")

# Plot the counts
ggplot(sentiment_counts, aes(x = Sentiment, y = Count, fill = Sentiment)) +
  geom_bar(stat = "identity") +
  labs(title = "Counts of Positive, Negative, and Neutral Sentiments",
       x = "Sentiment Type",
       y = "Count") +
  theme_minimal()

```
In examining the sentiments uncover some intriguing insights.
The average positive sentiment score is around 0.066, indicating a modest level
of positivity in the narratives. In comparison, the mean negative sentiment
score is slightly lower at 0.042, suggesting that negativity is less prevalent 
in these tales., which is surprising to me. 

What stands out is the high neutral score of 0.892. This reflects a tendency 
for the language in the stories to be descriptive and straightforward, rather 
than heavily emotional. The overall compound score, which averages 0.054, 
supports this observation, indicating that while there is a hint of positivity,
the overall sentiment leans more towards neutrality.

Overall, these findings provide a snapshot of the sentiment present in the
sample analyzed, but they may not be representative of every story within 
Grimm's Fairy Tales due to the limited nature of the sample.


**Reference**

Silge, Julia, and David Robinson. "Sentiment Analysis." Tidy Text Mining.
Last modified August 21, 2023. https://www.tidytextmining.com/sentiment#most-
positive-negative.

GeeksforGeeks. 2024. “Python Sentiment Analysis Using VADER.” 
Accessed November 2, 2024. https://www.geeksforgeeks.org/python-sentiment-
analysis-using-vader/.

“VADER: Valence Aware Dictionary and sEntiment Reasoner.” 2024. 
Accessed November 2, 2024. https://cran.r-project.org/web/packages/vader/vader.pdf.


*Chat gpt Questions*

What would be the best soultion in fixing Warning: number of items to replace 
is not a multiple of replacement length.

How to speed up VADER processing? 